Generative Adversarial Networks

### 2017-12-04 cs6735/project/report.tex
small updates (computer is crashing)

\begin{abstract}
    Within the past decade, deep generative networks have often fallen short, especially when compared to deep discriminative models. However, generative adversarial networks (GANs), first introduced in 2014 by Goodfollow \textit{et al.} \citep{2014arXiv1406.2661G}, have proved themselves as a promising research avenue for deep generative models. GANs present a new architecuture which allows for the unsupervised training of a generative model through backpropogation. Due to the current research effort, state-of-the-art results are now being acheived on a regular basis \citep{2017arXiv171010196K, 2017arXiv171111585W}; however, GANs still encounter instability during training and do not yet poccess the ability to produce completly realistic images. Nevertheless, these networks offer encouraging results for unsupervised learning, semi-supervised, reinforcement learning and several image generation tasks. This research paper will examine the motivations and challenges associated with GANs, their training process, architecture variations and some applications that are currently being explored.
\end{abstract}

@@

\subsection{Generative Adversarial Networks}
Generative adversarial networks (GANs), first introduced in 2014 by Goodfellow \textit{et al.}, offer a new framework for estimating generative models with use of an adversarial process \citep{2014arXiv1406.2661G}. These models offer a clever approach to solving the aforementioned problem of explicitly defining a cost function. Rather than training a single model, a discriminator is introduced. These two networks are pitted against each other in a minimax, zero-sum game. The generative model $G$ attempts to produce data that resembles that of the training set while the discriminative model $D$ attempts to discriminate between a real and fake (generated) sample. To train this network, $G$ attempts to maximize the error rate of $D$ whereas $D$ attempts to minimize it. The goal of the GAN algorithm is to reduce the distance between the real data and generated data probability distributions. Backpropagation is used to update the parameters of both the descriminator and generator. In the ideal case, $D$ is unable to distinguish generated samples from real samples and produces an error rate of $\frac{1}{2}$ \citep{2014arXiv1406.2661G}.

@@

\section{Background}
The goal of the GAN training process is to determine the discriminator and generator parameters such that the classifier error is minimized and maximized respectively \citep{2017arXiv171007035C}. GANs may use any differentiable functions for their underlying implementation; however, optimal results are typically achieved when neural networks are used \citep{2014arXiv1406.2661G}. We first define a vector $z$ which is sampled from a prior noise distribution $z~p_z(z)$. This vector becomes the argument of the generator $G(z;\theta_g)$ where $G$ is a differentiable function and $\theta_g$ are its parameters. Another differentiable function $D(x;\theta_d)$ is defined for the discriminator. This function takes as input a vector $x$ from either the output of $G$ or the training set and is parametrized by $\theta_d$. The output of the discriminator is a scalar number which represents the probability that the sample $x$ was derived from the training data. In this situation, there exists a minimax game, a scenario derived from game theory, where $D$ attempts to maximize the probability of assigning a correct label and $G$ attempts to minimize that same probability. The value function $V(D,G)$ becomes:

@@

\section{Issues}
There exist several issues that characterize GANs. Some are due to the architecture of GANs while others to implementations of the generator and descriminator. These issues lay at the forefront of GAN research.

\subsection{Instability} \label{sec:instability}
GANs are often characterized by their instability \citep{2017arXiv170100160G}. The training process makes use of gradient descent which does not guarantee convergence to the Nash equilibrium, the game theory equilibrium that involves two competing players. This training process for GANs is different than network optimization which is common in discriminative models. When either the generator or discriminator trains, even if that specific network improves, the other may regress. This sometimes leads to the scenario where each network repeatedly undo's the other's progress. This problem is not specific to GANs, but to any zero-sum game situations \citep{2017arXiv170100160G}.

Goodfellow has demonstrated that convergence is guaranteed if updates are made directly to the function space; however, this is obviously not what occurs. During training, we only ever update the parameter space \citep{2014arXiv1406.2661G}. As of yet, no theoretical justifications have been found that guarantee convergence or non-convergence when updates are only made to the parameters \citep{2017arXiv170100160G}. In practice, adversarial networks often oscillate between different kinds of samples without converging to an equilibrium in a phenomenon called mode collapse.

\subsection{Mode Collapse}
Mode collapse occurs when the generator collapses several different latent space vectors $z$ to the same outputs. This problem can be visualized in figure \ref{fig:modecollapse}. Full mode collapse is rare; however, partial mode collapse often occurs and varies by intensity \citep{2017arXiv170100160G}. For example, the generator may create several images that contain the same object at different perspectives, but generated with dissimilar vectors. This occurs when the capacity of the generator is superior to that of the discriminator. Backpropagation does not necessarily follow the minimax game mentioned earlier and may behave like the maximin version:

@@

When this occurs, the generator attempts to map every $z$ vector to the one $x$ variable the discriminator is most likely to guess is real. When training, one can only hope the behavior of gradient descent resembles the minimax version. It is often speculated that the root cause of the issue is the particular divergence measure used \citep{2017arXiv170100160G}. Specifically, it is often asserted that mode collapse is caused by the use of Jensen-Shannon divergence; however, the use of Kullback-Leibler divergence has be shown to worsen the problem \citep{2017arXiv170100160G}. As mode collapse is currently an unavoidable reality for GANs, their applications are limited to situations where it is acceptable to generate only a small number of distinct outputs \citep{2017arXiv170100160G}. However, recent progress has been made to reduce this problem. A potential solution is discussed in section \ref{sec:minidisc}.

\begin{figure}
\centering
\includegraphics[scale=0.8]{modecollapse}
\caption{An example of mode collapse as demonstrated by Metz \textit{et al.} \citep{2016arXiv161102163M}. The generated samples cluster around only one of the modes and do not spread out as they should. Throughout the training process, this cycle continues does not reach an equilibrium.}
\label{fig:modecollapse}
\end{figure}

>>>>

\begin{abstract}
    Within the past decade, deep generative networks have often fallen short, especially when compared to deep discriminative models. However, generative adversarial networks (GANs), first introduced in 2014 by Goodfellow \textit{et al.} \citep{2014arXiv1406.2661G}, have proved themselves as a promising research avenue for deep generative models. GANs present a new architecture which allows for the unsupervised training of a generative model through backpropogation. Due to the current research effort, state-of-the-art results are now being achieved on a regular basis \citep{2017arXiv171010196K, 2017arXiv171111585W}; however, GANs still encounter instability during training and do not yet poccess the ability to produce completely realistic images. Nevertheless, these networks offer encouraging results for unsupervised learning, semi-supervised, reinforcement learning and several image generation tasks. This research paper will examine the motivations and challenges associated with GANs, their training process, architecture variations and some applications that are currently being explored.
\end{abstract}

@@

\subsection{Generative Adversarial Networks}
Generative adversarial networks (GANs), first introduced in 2014 by Goodfellow \textit{et al.}, offer a new framework for estimating generative models with use of an adversarial process \citep{2014arXiv1406.2661G}. These models offer a clever approach to solving the aforementioned problem of explicitly defining a cost function. Rather than training a single model, a discriminator is introduced. These two networks are pitted against each other in a minimax, zero-sum game. The generative model $G$ attempts to produce data that resembles that of the training set while the discriminative model $D$ attempts to discriminate between a real and fake (generated) sample. To train this network, $G$ attempts to maximize the error rate of $D$ whereas $D$ attempts to minimize it. The goal of the GAN algorithm is to reduce the distance between the real data and generated data probability distributions. Backpropagation is used to update the parameters of both the discriminator and generator. In the ideal case, $D$ is unable to distinguish generated samples from real samples and produces an error rate of $\frac{1}{2}$ \citep{2014arXiv1406.2661G}.

@@

\section{Background} \label{sec:background}
The goal of the GAN training process is to determine the discriminator and generator parameters such that the classifier error is minimized and maximized respectively \citep{2017arXiv171007035C}. GANs may use any differentiable functions for their underlying implementation; however, optimal results are typically achieved when neural networks are used \citep{2014arXiv1406.2661G}. We first define a vector $z$ which is sampled from a prior noise distribution $z~p_z(z)$. This vector becomes the argument of the generator $G(z;\theta_g)$ where $G$ is a differentiable function and $\theta_g$ are its parameters. Another differentiable function $D(x;\theta_d)$ is defined for the discriminator. This function takes as input a vector $x$ from either the output of $G$ or the training set and is parametrized by $\theta_d$. The output of the discriminator is a scalar number which represents the probability that the sample $x$ was derived from the training data. In this situation, there exists a minimax game, a scenario derived from game theory, where $D$ attempts to maximize the probability of assigning a correct label and $G$ attempts to minimize that same probability. The value function $V(D,G)$ becomes:

@@

\section{Issues}
There exist several issues that characterize GANs. Some are due to their architecture while others to implementations of the generator and discriminator. These issues lay at the forefront of GAN research.

\subsection{Instability} \label{sec:instability}
GANs are often characterized by their instability \citep{2017arXiv170100160G}. The training process makes use of gradient descent which does not guarantee convergence to the Nash equilibrium, the game theory equilibrium that involves two competing players. This training process is different than network optimization which is common in discriminative models. When either the generator or discriminator trains, even if that specific network improves, the other may regress. This sometimes leads to the scenario where each network repeatedly undo's the other's progress. This problem is not specific to GANs, but to any zero-sum game situations \citep{2017arXiv170100160G}.

Goodfellow has demonstrated that convergence is guaranteed if updates are made directly to the function space; however, this is not what occurs in practice. During training, we make updates to the parameter space rather than the function space \citep{2014arXiv1406.2661G}. As of yet, no theoretical justifications have been found that guarantee convergence or non-convergence when updates are made to the parameters \citep{2017arXiv170100160G}. In practice, adversarial networks often oscillate between different kinds of samples without converging to an equilibrium in a phenomenon called mode collapse.

\subsection{Mode Collapse}
Mode collapse occurs when the generator collapses several different latent space vectors $z$ to the same outputs. This problem can be visualized in figure \ref{fig:modecollapse}. Full mode collapse is rare; however, partial mode collapse often occurs and varies by intensity \citep{2017arXiv170100160G}. For example, the generator may create several images that contain the same object at different perspectives, but generated with dissimilar vectors. This occurs when the capacity of the generator is superior to that of the discriminator. Backpropagation does not necessarily follow the minimax game mentioned is section \ref{sec:background} and may behave like the maximin version:

@@

When this occurs, the generator attempts to map every $z$ vector to the sample $x$ that which the discriminator is most likely to guess is real. When training, one can only hope that the behavior of gradient descent resembles that of the minimax game \citep{2017arXiv170100160G}. It is often speculated that the root cause of the issue is the particular divergence measure used \citep{2017arXiv170100160G}. Specifically, it is often asserted that mode collapse is caused by the use of Jensen-Shannon divergence; however, the use of Kullback-Leibler divergence has been shown to worsen the problem \citep{2017arXiv170100160G}. As partial mode collapse is currently an unavoidable reality, the applications of GANs are limited to situations where it is acceptable to generate only a small number of distinct outputs \citep{2017arXiv170100160G}. However, recent progress has been made to reduce this problem. A potential solution is discussed in section \ref{sec:minidisc}.

\begin{figure}
\centering
\includegraphics[scale=0.8]{modecollapse}
\caption{An example of mode collapse as demonstrated by Metz \textit{et al.} \citep{2016arXiv161102163M}. The generated samples cluster around only one of the modes and do not spread out as would be expected. Throughout the training process, this cycle continues and does not reach an equilibrium.}
\label{fig:modecollapse}
\end{figure}
